Supporting Information:\\ NMRlipids IV: Headgroup \& glycerol backbone structures, and cation binding in bilayers with PS lipids

### 2019-04-23 Manuscript/manuscriptPSsuppl.tex

>>>>

\pagebreak
\section{Sensitivity of the molecular electrometer to the chosen definition of ion concentration}
%\section{Effect of the definition of concentration on the headgroup order parameters as a function of ion concentration.}\label{concentrationDEFsection}
Previous studies using the electrometer concept to assess the ion binding affinity to lipid
bilayers report ion concentrations either in water before solvating the lipids (buffer concentration) \cite{akutsu81,roux90,catte16}
or in bulk water after solvating the lipids (bulk concentration) \cite{altenbach84,melcr18}.
In this work, we use the former definition to be consistent with the experimental reference data \cite{roux90}. The difference between these two concentrations increases
with the increasing ion binding affinity. However, Fig. \ref{concentrationDEFfigure} demonstrates that 
in a model with realistic ion binding affinity~\cite{melcr18} the deviation between the two definitions is neglible.
\begin{figure}[]
  \centering
  \includegraphics[width=8.3cm]{../Figs/OP_ECC_POPC_DPPC_water_conc2_dppc_bulk.pdf}
  \caption{\label{concentrationDEFfigure}
    Changes of the headgroup order parameters as a function of CaCl$_2$ concentration using
    two possible definitions of ion concentration from the recent force field
    with realistic calcium binding affinity to a POPC bilayer \cite{melcr18}
    together with the experimental data \cite{akutsu81,altenbach84}. 
  }
\end{figure}
